Geometric variational problems

### 2020-11-16 NSF2020/Description/background.tex
edited background.tex

\begin{wrapfigure}[10]{r}{0.22\textwidth}
\centerline{\includegraphics[width=0.22\textwidth]{figures/BG_fig1.pdf}}
\caption{ \footnotesize
A collection of rigid particles: $P_1,$ $P_2,$ $P_3, \ldots$. The
exterior domain $\Omega$ represents the solvent.}
  \label{fig:domain}
\end{wrapfigure}
Recently, PIs RR and YNY developed a mathematical model, called the
hydrophobic attraction potential (HAP) model \cite{Fu2018_SIAM}, that is based
on the physical origin of hydrophobicity. This model addresses the major
shortcomings of molecular dynamics (MD) and continuum approaches. Based
on preliminary results (\S\ref{sec:preliminary_work}), the PIs propose to
extend this HAP model to offer an alternative modeling methodology that
leads to new mathematical ideas and is both physically accurate and
computationally practical.
%
%
%The word hydrophobic (water fearing) derives from the low solubility of oil (hydrocarbon solute) in water and vice versa. 
%It causes hydrophobic moieties to aggregate and cluster,
%is responsible for the adhesion between hydrophobic surfaces \cite{Ducker2016}, large contact angles on a 
%dewetting surface \cite{Arenas2019,Sandre1999}, accumulation of particles along interfaces \cite{Lee2013,Lee2014}, 
%formation of micelles and bilayers \cite{Israelachvili80}, and protein folding and membrane insertion \cite{Kabelka2018}.
%
%Lipids are amphiphilic molecules whose  structure possesses
%both hydrophobic and hydrophilic parts. 
%The amphiphilic property is what allows lipids to form the membranes and 
%compartments of living cells \cite{Israelachvili80}. 
%More specifically, a lipid consists of 
%an elongated hydrocarbon tail that is hydrophobic, attached to a polar head that is hydrophilic.
%To shield the hydrophobic tails from water, lipids self-assemble into micelles and bilayers. 
%A micelle is a spherical arrangement of lipids with tails terminating at the micelle center. 
%A bilayer consists of two layers of lipids called monolayers, where the lipid tails point 
%from the monolayer surface into the bilayer core. 
%
%The mathematical modeling of a biological membrane is a challenging problem in applied mathematics. 
%Bilayers are elastic and resist deformations like bending, twisting, and stretching.
%Their elastic deformations are well described by the theory of liquid crystals \cite{ANDRIENKO2018520}.
%Lipid bilayer membranes can also be fluidic, and the lateral translation of lipids (or any membrane bound
%proteins) couples nonlocaly to the motion of the aqueous environment \cite{MerkelSackmannEvans1989,StoneAjdari1998_JFM,OppenheimerDiamant2009_BJ,OppenheimerDiamant2011_PRL}. 
%%Finally, the membranes of cellular 
%%compartments are constantly merging and pinching as part of intracellular trafficking. 
%%Therefore monolayer surfaces undergo discontinuous deformations. 
%
%There are two prevailing approaches in membrane modeling, and each has its advantages
%and disadvantages. Molecular dynamics (MD) is to date the only tool capable of resolving granular biological details
%The computational cost of MD, however, grows with the sixth power of the sample diameter and 
%so simulations are severely limited to 
%small system sizes and short time scales \cite{DiCarlo2019}.
%The other approach, continuum mechanics, assumes smooth surfaces and can therefore model 
%realistic systems over physical times. 
%But continuum description of a biological membrane ignores the granularity of lipid molecules, and thus requires some assumptions when a
%membrane ruptures or when two membranes fuse \cite{ChKo08}. 
%%
%%mechanics presumes, rather than predicts, the nucleation of discontinuities. 
%%
%
The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles. These results will facilitate optimal design of smart materials by
tuning the geometry and properties of the amphiphilic particles.

\subsection{Hydrophobic Attraction Potential (HAP)}
\label{sec:HAP}
%Based on the physical origin of hydrophobicity,  we have devised a
%functional, called the hydrophobic attraction potential (HAP), to model
%hydrophobic forces~\cite{Fu2018_SIAM}. 
The motivation for the HAP concept stemmed from PI RR's work concerning
energy barriers in membrane
fusion~\cite{RyKlYaCo16,Chetal16}. By applying a mathematical, squared-gradient
theory for hydrophobic attraction between planar
surfaces~\cite{Eriksson1989,Lum1999,Menshikov2017,Marcelja1977}, PI RR
and collaborators resolved the long-standing issue of accounting for the
energy of a monolayer fissure surface during topological transitions.
Based on PI RR and YNY's coarse-grained membrane modeling
work~\cite{Fu2017}, the investigators devised a gradient theory for
arbitrary collections of hydrophobic and
amphiphilic particles. As a new method, HAP eliminates the costly calculation
of water by treating the solvent implicitly, is capable of representing
biologically-relevant morphologies, and avoids complicated
re-meshing schemes of continuum approaches by utilizing a particle-based
representation.

@@

\boldsymbol{\sigma}_{\text{hydro}} = \gamma \rho^{-1} u^2 I + 2\rho \gamma (\tfrac{1}{2}|\nabla u|^2I - \nabla u \nabla u^T).

@@

The time integration of
particle configurations goes as follows: \textbf{(i)} solve the
BVP~\eqref{SL} for the screened Laplace equation, \textbf{(ii)}
determine the rigid body forces and torques~\eqref{forceandtorque},
\textbf{(iii)} solve the Stokes mobility problem~\eqref{eq:stokes} for
rigid body motion, and \textbf{(iv)} update the particle configuration.
Numerical challenges and how these challenges are overcome are described
in \S\ref{subsec:specific_aim_2}.

>>>>

Recently, PIs RR and YNY developed a mathematical model, called the
hydrophobic attraction potential (HAP) model \cite{Fu2018_SIAM}, that is based
on the physical origin of hydrophobicity. This model addresses the major
shortcomings of molecular dynamics (MD) and continuum approaches. Based
on preliminary results (\S\ref{sec:preliminary_work}), the PIs propose to
extend this HAP model to offer an alternative modeling methodology that
leads to new mathematical ideas and is both physically accurate and
computationally practical.
%
%
%The word hydrophobic (water fearing) derives from the low solubility of oil (hydrocarbon solute) in water and vice versa. 
%It causes hydrophobic moieties to aggregate and cluster,
%is responsible for the adhesion between hydrophobic surfaces \cite{Ducker2016}, large contact angles on a 
%dewetting surface \cite{Arenas2019,Sandre1999}, accumulation of particles along interfaces \cite{Lee2013,Lee2014}, 
%formation of micelles and bilayers \cite{Israelachvili80}, and protein folding and membrane insertion \cite{Kabelka2018}.
%
%Lipids are amphiphilic molecules whose  structure possesses
%both hydrophobic and hydrophilic parts. 
%The amphiphilic property is what allows lipids to form the membranes and 
%compartments of living cells \cite{Israelachvili80}. 
%More specifically, a lipid consists of 
%an elongated hydrocarbon tail that is hydrophobic, attached to a polar head that is hydrophilic.
%To shield the hydrophobic tails from water, lipids self-assemble into micelles and bilayers. 
%A micelle is a spherical arrangement of lipids with tails terminating at the micelle center. 
%A bilayer consists of two layers of lipids called monolayers, where the lipid tails point 
%from the monolayer surface into the bilayer core. 
%
%The mathematical modeling of a biological membrane is a challenging problem in applied mathematics. 
%Bilayers are elastic and resist deformations like bending, twisting, and stretching.
%Their elastic deformations are well described by the theory of liquid crystals \cite{ANDRIENKO2018520}.
%Lipid bilayer membranes can also be fluidic, and the lateral translation of lipids (or any membrane bound
%proteins) couples nonlocaly to the motion of the aqueous environment \cite{MerkelSackmannEvans1989,StoneAjdari1998_JFM,OppenheimerDiamant2009_BJ,OppenheimerDiamant2011_PRL}. 
%%Finally, the membranes of cellular 
%%compartments are constantly merging and pinching as part of intracellular trafficking. 
%%Therefore monolayer surfaces undergo discontinuous deformations. 
%
%There are two prevailing approaches in membrane modeling, and each has its advantages
%and disadvantages. Molecular dynamics (MD) is to date the only tool capable of resolving granular biological details
%The computational cost of MD, however, grows with the sixth power of the sample diameter and 
%so simulations are severely limited to 
%small system sizes and short time scales \cite{DiCarlo2019}.
%The other approach, continuum mechanics, assumes smooth surfaces and can therefore model 
%realistic systems over physical times. 
%But continuum description of a biological membrane ignores the granularity of lipid molecules, and thus requires some assumptions when a
%membrane ruptures or when two membranes fuse \cite{ChKo08}. 
%%
%%mechanics presumes, rather than predicts, the nucleation of discontinuities. 
%%
%
The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles. These results will facilitate optimal design of smart materials by
tuning the geometry and properties of the amphiphilic particles.

\begin{wrapfigure}[9]{r}{0.24\textwidth}
  \vspace{-20pt}
\centerline{\includegraphics[width=0.24\textwidth]{figures/BG_fig1.pdf}}
\caption{ \footnotesize
A collection of rigid particles: $P_1,$ $P_2,$ $P_3, \ldots$. The
exterior domain $\Omega$ represents the solvent.}
  \label{fig:domain}
\end{wrapfigure}
\subsection{Hydrophobic Attraction Potential (HAP)}
\label{sec:HAP}
%Based on the physical origin of hydrophobicity,  we have devised a
%functional, called the hydrophobic attraction potential (HAP), to model
%hydrophobic forces~\cite{Fu2018_SIAM}. 
The motivation for the HAP concept stemmed from PI RR's work concerning
energy barriers in membrane
fusion~\cite{RyKlYaCo16,Chetal16}. By applying a mathematical, squared-gradient
theory for hydrophobic attraction between planar
surfaces~\cite{Eriksson1989,Lum1999,Menshikov2017,Marcelja1977}, PI RR
and collaborators resolved the long-standing issue of accounting for the
energy of a monolayer fissure surface during topological transitions.
Based on PI RR and YNY's coarse-grained membrane modeling
work~\cite{Fu2017}, the investigators devised a gradient theory for
arbitrary collections of hydrophobic and
amphiphilic particles. As a new method, HAP eliminates the costly calculation
of water by treating the solvent implicitly, is capable of representing
biologically-relevant morphologies, and avoids complicated
re-meshing schemes of continuum approaches by utilizing a particle-based
representation.

@@

\boldsymbol{\sigma}_{\text{hydro}} = \gamma \rho^{-1} u^2 I + 2\rho
  \gamma \left(\frac{1}{2}|\nabla u|^2I - \nabla u \nabla u^T\right).

@@

The time integration of particle configurations goes as follows:
\textbf{(i)} solve the BVP~\eqref{SL} for the screened Laplace equation,
\textbf{(ii)} determine the rigid body forces and
torques~\eqref{forceandtorque}, \textbf{(iii)} solve the Stokes mobility
problem~\eqref{eq:stokes} for the rigid body motions, and \textbf{(iv)}
update the particle configuration.  Numerical challenges and how these
challenges are overcome are described in \S\ref{subsec:specific_aim_2}.
